Robust Control of a Fluxonium Qubit

### 2020-09-25 paper/main.tex
updates to figs and nbs

\documentclass[
  amsfonts,
  amsmath,
  tbtags,
  amssymb,
  aps,
  nobibnotes,
  %% prl,
  twocolumn,
]{revtex4-2}

@@

%% TITLE
\title{Robust Quantum Optimal Control}

\author{Thomas Propson}
\email{tcpropson@uchicago.edu}
\affiliation{
  Department of Physics, University of Chicago, Chicago, Illinois 60637, USA
}

@@

%% ABSTRACT
\begin{abstract}
  %% - 1 + 2 topic importance
  %% - 1 what I have done
  %% - few sentences about primary results
  The ability to engineer high fidelity operations on quantum processors in the presence of
  calibration errors and decoherence remains the primary challenge requisite to achieving quantum advantage.
  Quantum optimal control (QOC) techniques have proven effective in realizing high fidelity operations,
  but they require exquisite calibration to be performant. In this work we employ robust control techniques
  to achieve gate errors two orders of magnitude lower than dynamic decoupling for parameter
  spreads on the order of $1\%$. 
  Additionally, current QOC techniques mitigate decoherence effects due to longitudinal relaxation by integrating
  Lindblad master equation dynamics, which exponentially increase the computational complexity
  of the optimization algorithm.
  In this work we achieve a factor of $5$ increase in longitudinal relaxation times
  for the gate set of a fluxonium qubit
  using a computationally efficient metric in optimizaiton.
\end{abstract}

@@

%% S1
\section{Introduction}
textwidth: \printinunitsof{in}\prntlen{\textwidth}
linewidth: \printinunitsof{in}\prntlen{\linewidth}
%% P1 - general field, lots of citation. cite reviews, original qoc.
%% P2 - exhaustively find all papers, what people have done to address the issues
%% P3 - what we do is. maybe add a rant about how
%% QOC is classical control theory but we've ignored that field
%% P4 - outline
(The field) The field of quantum optimal control (QOC) is concerned
with improving the efficiency and accuracy with which quantum systems
are manipulated.
Early QOC techniques were proposed for nuclear magnetic resonance experiments
\cite{khaneja2005optimal}, and applications now include superconducting
qubits \cite{heeres2017implementing,
  leng2019robust, leung2017speedup, xu2020nonadiabatic},
neutral and ionized atoms \cite{van2016optimal}, nitrogen-vacancy centers in
diamond \cite{rembold2020introduction}, and Bose-Einstien condensates
\cite{sorensen2018quantum}.
In quantum control theory, as in classical control theory,
optimization is performed
to minimize an objective function. Relevant for the field
of quantum computation is the task of achieving high fidelity quantum gates.
In this application, the objective function includes contributions from the
gate error associated with the state transfer of the system as well as contributions
due to experimental constraints, such as the control amplitude.
The decision variables of the optimization problem are the time-dependent control
parameters, particular to the quantum system, that govern its evolution.

@@

(This work) In this work, we employ state of the art
trajectory optimization methods
to formulate the quantum
optimal control problem as a constrained optimization problem, allowing
us to enforce experimental constraints.
Additionally,
we introduce methods for constructing gates that mitigate longitudinal
relaxation and are robust to experimental errors, and therefore
pure dephasing, the dominant barriers to experimentally
realizing practical quantum computing.

(Outline) As an example, we study the quantum optimal control problem on
the fluxonium qubit presented in \cite{zhang2020universal}.
We describe experimentally realistic constraints and map them to
the trajectory optimization framework. Next, we
outline a method for making the optimization longitudinal
relaxation aware. We achieve a 4x increase in longitudinal relaxation times
over the baseline gate set.
We present two methods for achieving robustness to pure dephasing
and compare them to existing dynamic decoupling methods. We find that
the methods we employ outperform dynamic decoupling, producing
high fidelity gates at system parameter spreads of $5\%$.

@@

%% F1
\begin{figure}[ht]
  \begin{subfigure}{.55\linewidth}
    \includegraphics[width=\linewidth]{assets/f1a.png}
  \end{subfigure}\hfill
  \begin{subfigure}{.4\linewidth}
    \includegraphics[width=\linewidth]{assets/f1b.png}
  \end{subfigure}

  \begin{subfigure}{\linewidth}
    \includegraphics[width=\linewidth]{assets/f1c.png}
  \end{subfigure}
  \caption{
    (a) $T_{1}$ optimized gates (red) and analytic gates (blue).
    (b) $T_{1}$ interpolation function used in optimization. Markers
    denote the time-averaged, absolute amplitude of each gate.
    (c) Master equation simulation with $T_{1}$ dissipation
    over the course of concatenated gate applications.
  }
\end{figure}

%% S3
\section{QOC on the Fluxonium}
(Fluxonium Device) In the following we study
the quantum optimal control problem on the fluxonium qubit.
In the two-level
approximation the system Hamiltonian takes the form
\label{eq:hamiltonian}
\begin{align}
  H/h &= f_{q} \frac{\sigma_{z}}{2} + a(t) \frac{\sigma_{x}}{2}
\end{align}
where $f_{q}$ is the qubit frequency at the flux frustration point,
$a$ is the flux drive amplitude, $h$ is Planck's constant, and $\sigma_{x}, \sigma_{y}$
are Pauli matrices. The flux amplitude $a$ is experimentally
realized by modulating the flux 
threading the device ($\Phi_{\textrm{ext}}$).
The flux amplitude is obtained from the external flux by
$a = 4 \pi \braket{g \lvert \hat{\phi} \rvert e} \rvert_{0.5 \Phi_{0}} E_{L}
\delta \Phi_{\textrm{ext}} / (h \Phi_{0})$
where $\hat{\phi}$ is the phase operator, $E_{L}$ is the characteristic inductance energy, $\Phi_{0}$
is a flux quantum, and
$\delta \Phi_{\textrm{ext}} = \Phi_{\textrm{ext}} - 0.5 \Phi_{0}$ is the flux
offset from the flux frustration point.

In the following we compare the gates found with our methods
to the $X/2, Y/2, Z/2$ gates reported in \cite{zhang2020universal}
for the same device. Note that $Y/2$ and arbitrary $Z$ rotations
are sufficient for universal computation.
%% TODO: list device parameters

(Constraints) We now outline the constraints that we require each gate
to obey.
We require $a(t = 0) = a(t = t_{N}) = 0$.
This constraint ensures gates may be concatenated arbitrarily without
inducing AWG ringing due to high-frequency transitions.
Furthermore, we require $\int_{0}^{t_{N}} a(t) dt = 0$. This
constraint ensures the pulse has zero net flux, mitigating
the hysteresis ubiquitous in flux bias lines.
%% TODO: See ref. 28 in Helin paper
We require $-300 \textrm{MHz} \le a(t) \le 300 \textrm{MHz}$
to ensure the two-level approximation \ref{eq:hamiltonian}
remains valid. Additionally, we require that each gate achieves
the desired state transition $\braket{\psi_{f} \lvert \psi_{N}} = 1$.
In addition to these constraints we penalize the norm
of the first and second derivatives of the flux amplitude
to mitigate AWG ringing.

The optimization is performed over the second derivative of the flux amplitude
($\frac{d^{2}}{dt^{2}} a(t)$) which is contained in the
augmented control vector. The first derivative
($\frac{d}{dt} a(t)$), proportional ($a(t)$), and integral ($\int a(t)$)
flux amplitude terms
are contained in the augmented state vector. They are obtained from
the second derivative of the flux amplitude by
integration in the dynamics function $f$.
Both the zero net flux and target quantum state constraint
are then handled by ensuring the augmented final state is
reached $x_{N} - x_{f} = 0$.
The equality and inequality constraints on $a(t)$ are handled
with a bound constraint.
%% TODO: probably want to list equations for the constraints

($T_{1}$ and $T_{\phi}$ noise)
In addition to optimizing the cost functional to achieve a gate
that obeys experimental constraints and has a high simulated fidelity,
we also want to make the gate robust to noise that affects the experimental
gate fidelity. Decoherence of the quantum state due to external noise
is typically modeled by two phenomena: longitudinal relaxation and pure dephasing.
They are modeled using their $1/e$ decay times $T_{1}$ and $T_{\phi}$ respectively
(see Appendix).
The main contributions to longitudinal relaxation in our
device are dielectric loss in the capacitor, resistive loss in the inductor,
and Purcell loss. The main contributions to pure dephasing in our
device are $1/f$ flux noise and decay via charge and flux coupling
to the control lines.

Dissipation to the thermal bath via longitudinal
relaxation is an irreversible process
that results in information loss.
Converesely, pure dephasing is a reversible process.
There is a tradeoff between the two decoherence processes. In the case of white
noise we have that the sum of the noise weights $W_{1}$ and $W_{\phi}$
is constant \cite{huang2020engineering}.
Our device achieves its best pure dephasing
protection at the flux frustration point
$T_{2e}(\Phi_{\textrm{ext}} = 0.5 \Phi_{0}) \sim 300 \mu\textrm{s}$
where the qubit frequency is first-order insensitive to changes in flux.
It becomes more succeptable to pure dephasing as the flux is tuned away from the flux
frustration point. Conversely, $T_{1}$ is at a minimum
at the flux frustration point $T_{1}(0.5 \Phi_{0}) = 0.315$ms,
and increases away from the flux frustration point
$T_{1}(0.43 \Phi_{0}) = 4.3$ms. Given the nature
of the decay processes and the tradeoff, we choose
to maximize the longitudinal relaxation time ($T_{1}$)
over the gate duration
and employ robust control techniques to mitigate
pure dephasing.
%% TODO: See fig. 3 in Helin paper

%% S4
\section{Longitudinal Relaxation Awareness}
(Strategy) We seek to minimize the probability
that the qubit decays as a result of longitudinal
relaxation. To this end we add the
longitudinal relaxation probability
to the augmented state vector
\begin{equation}
  P_{1}(t) = \int_{0}^{t} \gamma_{1}(a(t^{\prime})) dt^{\prime}
\end{equation}
where $\gamma_{1} = T_{1}^{-1}$. Setting the target
longitudinal relaxation probability to 0 results in
a quadratic cost at each knot point
of the form ${\lvert P_{1}(t_{k}) \rvert}^{2}$.

$\gamma_{1}(a_{k})$
is obtained at each knot point by evaluating
a spline interpolant fit to
experimentally obtained data of the form
$\{(\Phi_{\textrm{ext}}, T_{1})\}$.
$\Phi_{\textrm{ext}}$ is a function of $a_{k}$
with the inverse of the relation given in section 3.
Calculating $T_{1}$ directly from theoretical
considerations requires many high-dimensional
eigendecompositions, which
is computationally expensive. Additionally,
$T_{1}$ values are known to fluctuate greatly
with laboratory temperatures \cite{klimov2018fluctuations}.
Interpolating $T_{1}$ from experimental data increases
the fridge truth of the simulation.

Furthermore, the probability of longitudinal 
relaxation is dependent on the gate duration $t_{N}$.
We allow the optimizer to tune the gate duration by
adding the time step between knot points $\Delta t_{k}$
to the augmented control vector $u_{k}$.
Promoting $\Delta t_{k}$ to a decision variable, rather
than the number of knot points $N$, preserves the
Markovian decision structure of the trajectory
optimization problem. To ensure numerical
integration accuracy is maintained we add a bound
constraint at each knot point
$5\textrm{e-}3 \ \textrm{ns} \le
\Delta t_{k} \le 2\textrm{e-}2 \ \textrm{ns}$.
Note that this bound constraint is allowed to be
broken for intermediate iterations of the optimization,
so we use the absolute value of the time step
$\lvert \Delta t_{k} \rvert$ to ensure that it is non-negative.

(Results) We achieve a factor of 5 decrease in the probability
of longitudinal relaxation from the analytic gates we benchmark
against. We simulate the performance of the gates using the
Lindblad master equation (see Appendix). We repeatedly apply the basis gates
and measure the fidelity of the resulting state as a function
of time as shown in Fig. 1.

%% T1
%% TODO: maybe for appendix.
\begin{table}[ht]
  \begin{tabular}{c | c | c | c}
    & Analytic  & QOC &\\
    Gate & $P_{1}\ (10^{-5})$ & $P_{1}\ (10^{-5})$ & $P_{1\textrm{A}} / P_{1\textrm{Q}}$\\
    \hline
    Z/2 & 5.745 & 1.149 & 5.000\\
    Y/2 & 5.253 & 1.157 & 4.540\\
    X/2 & 16.251 & 6.946 & 2.340\\
  \end{tabular}
  \caption{Probability of longitudinal relaxation for each gate
    evaluated at the gate's duration.}
\end{table}

%% F2
\begin{figure}[ht]
  \begin{subfigure}{.55\linewidth}
    \includegraphics[width=\linewidth]{assets/f2a.png}
  \end{subfigure}\hfill
  \begin{subfigure}{.4\linewidth}
    \includegraphics[width=\linewidth]{assets/f2b.png}
  \end{subfigure}

  \begin{subfigure}{\linewidth}
    \includegraphics[width=\linewidth]{assets/f2c.png}
  \end{subfigure}

  \caption{
    (a) X/2 gates robust to qubit frequency detunings constructed using the
    analytic, (2, 4)-point sampling, and (2\textsuperscript{nd},
    3\textsuperscript{rd})-order derivative methods.
    (b) Gate error as a function of the detuning from the nominal
    qubit frequency at a fixed gate duration, see (a), and at the best
    gate duration for each method chosen from a sweep, see (c).
    (c) Gate error at a one-percent detuning from the nominal qubit frequency as
    a function of the gate duration set in optimization.
  }
\end{figure}

%% F3
\begin{figure}[ht]
  \begin{subfigure}{\linewidth}
    \includegraphics[width=\linewidth]{assets/f3a.png}
  \end{subfigure}

@@

  %% \begin{subfigure}{.5\linewidth}
  %%   \includegraphics[width=\linewidth]{assets/f3c.png}
  %% \end{subfigure}

  \caption{
    (a) $X/2$ gates robust to flux offsets constructed using the analytic,
    (2, 4)-point sampling, and (2, 3)-derivative methods at a gate duration
    of $56.8 (ns)$. (b) Mean gate error for 10 samples of 1/$f$ flux noise over 5000
    gate applications.
    %% (c) Stochastic 1/$f$ flux noise.
  }
\end{figure}

%% S5
\section{Robustness to Parameter Deviations}
(Strategy) For the fluxonium, pure dephasing acts
along the flux axis. Low frequency 1/$f$ noise,
Johnson-Nyquist current noise present in the flux bias
lines, and temperature dependent gain fluctuations
may cause the flux amplitude to deviate from its
desired value by an amount $\delta a$. Additionally,
residual calibration errors and fluctuations may cause
the qubit frequency to deviate from its measured value
by an amount $\delta f_{q}$.

System parameter
deviations are typically addressed with dynamic decoupling
sequences \cite{merrill2014progress},
the DRAG scheme \cite{krantz2019quantum}, or
geometric phase considerations
\cite{xu2020nonadiabatic} \cite{han2020experimental}.
Dynamic decoupling sequences compose rotations on the
Bloch sphere so that erroneous rotations arising due
to the parameter deviation are cancelled.
The latter two methods are inflexible. We draw on the
robust control literature to demonstrate
numerical techniques for engineering robustness
to parameter deviations.

We propose two methods for engineering robustness
to parameter deviations. The first we call the
derivative method. We draw on the intuition that
making the state evolution insensitive to changes
in the parameter ($x$) is encoded in the derivative of
the state with respect to that parameter
$\partial_{x}^{l} \ket{\psi}$.
The derviatives of the state are propgated in the
augmented state vector. Their dynamics are found by differentiating
the TDSE dynamics with respect to the parameter of interest
(see Appendix). Setting the target derivatives of the state
to zero vectors results in quadratic costs at each
knot point
${\lvert \partial_{x}^{l} \ket{\psi_{k}} \rvert}^{2}$.

The second scheme we analyze is the sampling method,
which is well studied in the robust control literature
\cite{manchester2016derivative} \cite{tronarp2016sigma}.
We add additional
states that follow deviant dynamics to the augmented
state vector. The parameters of interest
are altered by a small deviation, typically chosen
to represent the distribution of the parameter.
For example, in the 2-point sampling method we
propagate $\ket{\psi^{\pm, I}(t)}$ where
$f_{q}^{\pm} = f_{q} \pm \sigma_{f_{q}}$,
$f_{q}^{I} = f_{q}$.
The difference between the final states and the target
state is penalized, resulting in quadratic costs at
each knot point
${\lvert \ket{\psi^{\pm, I}_{k}} - \ket{\psi_{f}} \rvert}^{2}$

(Results) We compare the robust control
schemes to a dynamic decoupling pulse
for a single $R_{x}(\pi/2)$ gate
subject to qubit frequency detuning
($f_{q} \gets f_{q} + \delta f_{q}$)
to demonstrate the applicability of our techniques to mitigate
system parameter deviations.
We use a compensating for off-resonance error
(CORPSE) dynamic decoupling pulse that is optimized
to mitigate first-order error arising due to the detuning.
The pulse is composed of
three rotations about the $\hat{x}$, $-\hat{x}$, and $\hat{x}$ axes
of the Bloch sphere
designed to compensate for erroneous rotations
(see Appendix). We simulate the gate fidelity as
a function of the qubit frequency detuning for
variants of the
derivative method, sample method, and
CORPSE pulse (see Figure 2a). We also find that
increasing the gate duration allows the optimizer
to find a more robust pulse
(see Figure 2b). Increasing the order of the derivative
and sample methods requires propagating an additional quantum state
in the augmented state vector, a linear increase
in the computational complexity, but finding higher order
variants of the CORPSE family is prohibitively difficult.

>>>>

\documentclass[
  amsfonts,
  amsmath,
  tbtags,
  amssymb,
  aps,
  nobibnotes,
  %% prl,
  twocolumn,
  superscriptaddress,
]{revtex4-2}

@@

%% TITLE
\title{Robust Control of a Fluxonium Qubit}

\author{Thomas Propson}
\email{tcpropson@uchicago.edu}
\affiliation{
  James Franck Institute, University of Chicago, Chicago, Illinois 60637, USA
}
\affiliation{
  Department of Physics, University of Chicago, Chicago, Illinois 60637, USA
}
\author{Brian Jackson}
\author{Zac Manchester}
\affiliation{
  Department of Aeronautics and Astronautics Engineering, Stanford University, 496 Lomita Mall, Stanford, CA 94305
}
\author{David I. Schuster}
\affiliation{
  James Franck Institute, University of Chicago, Chicago, Illinois 60637, USA
}
\affiliation{
  Department of Physics, University of Chicago, Chicago, Illinois 60637, USA
}
\affiliation{
  Pritzker School of Molecular Engineering, University of Chicago, Chicago, Illinois 60637, USA
}

@@

%% ABSTRACT
\begin{abstract}
  %% - 1 + 2 topic importance
  %% - 1 what I have done
  %% - few sentences about primary results
  The ability to engineer high fidelity operations on quantum processors in the presence of
  calibration errors and decoherence remains the primary challenge requisite to achieving quantum advantage.
  Quantum optimal control (QOC) techniques have proven effective in realizing high fidelity operations,
  but they require exquisite calibration to be performant. In this work we employ robust trajectory optimization techniques
  to achieve high fidelity gates for parameter spreads on the order of $1\%$. 
  Additionally, current QOC techniques mitigate decoherence effects due to longitudinal relaxation by integrating
  the Lindblad master equation, which increases the computational complexity
  of the optimization algorithm.
  In this work we achieve a factor of $5$ increase in longitudinal relaxation times
  for the gate set of a fluxonium qubit
  using a computationally efficient metric in optimizaiton.
\end{abstract}

@@

%% S1
\section{Introduction}
textwidth: \printinunitsof{in}\prntlen{\textwidth}
linewidth: \printinunitsof{in}\prntlen{\linewidth}
%% P1 - general field, lots of citation. cite reviews, original qoc.
%% P2 - exhaustively find all papers, what people have done to address the issues
%% P3 - what we do is. maybe add a rant about how
%% QOC is classical control theory but we've ignored that field
%% P4 - outline
(The field) The field of quantum optimal control (QOC) is concerned
with efficiently and accurately manipulating quantum systems.
Early QOC techniques were proposed for nuclear magnetic resonance experiments
\cite{khaneja2005optimal}, and applications now include superconducting
qubits \cite{heeres2017implementing,
  leng2019robust, leung2017speedup, xu2020nonadiabatic},
neutral and ionized atoms \cite{van2016optimal}, nitrogen-vacancy centers in
diamond \cite{rembold2020introduction}, and Bose-Einstien condensates
\cite{sorensen2018quantum}.
In quantum control theory, as in classical control theory,
optimization is performed
to minimize an objective function. Relevant for the field
of quantum computation is the task of achieving high fidelity quantum gates.
In this application, the objective function includes contributions from the
gate error associated with the state transfer of the system as well as contributions
due to experimental constraints, such as the control amplitude.
The decision variables of the optimization problem are the time-dependent control
parameters, particular to the quantum system, that govern its evolution.

@@

(This work) In this work, we employ state of the art
trajectory optimization methods
to formulate the quantum
optimal control problem as a constrained optimization problem, allowing
us to enforce arbitrary experimental constraints.
Additionally,
we introduce methods for constructing gates that mitigate longitudinal
relaxation and are robust to experimental errors, and therefore
pure dephasing, the dominant barriers to experimentally
realizing practical quantum computing.

(Outline) As an example, we study the quantum optimal control problem on
the fluxonium qubit presented in \cite{zhang2020universal}.
We describe experimentally realistic constraints and map them to
the trajectory optimization framework. Next, we
outline a method for making the optimization longitudinal
relaxation aware. We achieve a factor of 5 increase in longitudinal relaxation times
over the baseline gate set.
We present two methods for achieving robustness to pure dephasing
and compare them to existing dynamic decoupling methods. We find that
the methods we employ are able to produce high fidelity gates when
subject to parameter spreads of $1\%$.

@@

%% S3
\section{QOC on the Fluxonium}
(Fluxonium Device) In the following we study
the quantum optimal control problem on the fluxonium qubit.
In the two-level
approximation the system Hamiltonian takes the form
\label{eq:hamiltonian}
\begin{align}
  H/h &= f_{q} \frac{\sigma_{z}}{2} + a(t) \frac{\sigma_{x}}{2}
\end{align}
where $f_{q} = 14$MHz is the qubit frequency at the flux frustration point,
$a$ is the flux drive amplitude, $h$ is Planck's constant, and $\sigma_{x}, \sigma_{y}$
are Pauli matrices. The flux amplitude $a$ is experimentally
realized by modulating the flux 
threading the device. We consider the task of constructing $Z/2$, $Y/2$, and $X/2$
gates for the fluxonium qubit subject to experimental constraints, decoherence, and
system parameter deviations. We compare the gates we obtain with optimal
control methods to the analytically constructed gates reported in
\cite{zhang2020universal} for the same device.
%% The flux amplitude is obtained from the external flux by
%% $a = 4 \pi \braket{g \lvert \hat{\phi} \rvert e} \rvert_{0.5 \Phi_{0}} E_{L}
%% \delta \Phi_{\textrm{ext}} / (h \Phi_{0})$
%% where $\hat{\phi}$ is the phase operator, $E_{L}$ is the characteristic inductance energy, $\Phi_{0}$
%% is a flux quantum, and
%% $\delta \Phi_{\textrm{ext}} = \Phi_{\textrm{ext}} - 0.5 \Phi_{0}$ is the flux
%% offset from the flux frustration point.

(Constraints) We now present the constraints we require each gate
to obey.
To ensures gates may be concatenated arbitrarily without
inducing AWG ringing due to high-frequency transitions,
we require $a(t = 0) = a(t = t_{N}) = 0$.
Furthermore, we require $\int_{0}^{t_{N}} a(t) dt = 0$. This
constraint ensures the pulse has zero net flux, mitigating
the hysteresis ubiquitous in flux bias lines.
%% TODO: See ref. 28 in Helin paper
We require $-0.5 \textrm{GHz} \le a(t) \le 0.5 \textrm{GHz}$
to ensure the two-level approximation \ref{eq:hamiltonian}
remains valid. Additionally, we require that each gate achieves
the desired state transition $\ket{\psi_{N}} = \ket{\psi_{f}}$.
In addition to these constraints we penalize the norm
of the first and second derivatives of the flux amplitude to
ensure its smoothness, and in doing so mitigate AWG ringing.

The optimization is performed over the second derivative of the flux amplitude
$\frac{d^{2}}{dt^{2}} a$ which is contained in the
augmented control vector. The first derivative
$\frac{d}{dt} a$, proportional $a$, and integral $\int a$
flux amplitude terms
are contained in the augmented state vector. They are obtained from
the second derivative of the flux amplitude by
integration in the dynamics function.
Both the zero net flux and target quantum state constraint
are then handled by ensuring the target augmented state is
reached $x_{N} = x_{f}$.
The equality and inequality constraints on $a$ are handled
with a bound constraint.
%% TODO: probably want to list equations for the constraints

($T_{1}$ and $T_{\phi}$ noise)
In addition to constructing gates
that obey experimental constraints and have high simulated fidelities,
we also desire to make the gates robust to noise that affects the experimental
gate fidelity. Decoherence of the quantum state due to external noise
is typically modeled by two phenomena: longitudinal relaxation and pure dephasing.
They are modeled using their $1/e$ decay times $T_{1}$ and $T_{\phi}$ respectively
(see Appendix).
The main contributions to longitudinal relaxation in our
device are dielectric loss in the capacitor, resistive loss in the inductor,
and Purcell loss. The main contributions to pure dephasing in our
device are $1/f$ flux noise and decay via charge and flux coupling
to the control lines.

Dissipation to the thermal bath via longitudinal
relaxation is an irreversible process
that results in information loss.
Converesely, pure dephasing is a reversible process.
There is a tradeoff between the two decoherence processes. In the case of white
noise we have that the sum of the noise weights $W_{1}$ and $W_{\phi}$
is constant \cite{huang2020engineering}.
Our device achieves its best pure dephasing
protection at the flux frustration point
$T_{2e}(a = 0) \sim 300 \mu\textrm{s}$
where the qubit is first-order insensitive to changes in flux.
It becomes more succeptable to pure dephasing as the flux is tuned away from the flux
frustration point $|a| > 0$. Conversely, longitudinal relaxation is at a minimum
at the flux frustration point $T_{1}(a = 0) = 0.315$ms,
and increases away from the flux frustration point
$T_{1}(a = 0.34 \textrm{GHz}) = 4.3$ms. Given the nature
of the decay processes and the tradeoff, we choose
to maximize the longitudinal relaxation time directly
via the flux amplitude and employ robust control techniques to mitigate
pure dephasing.

%% F1
\begin{figure*}[ht]
  \begin{subfigure}{.315\textwidth}
    \includegraphics[width=\linewidth]{assets/f1a.png}
  \end{subfigure}\hfill
  \begin{subfigure}{.23\textwidth}
    \includegraphics[width=\linewidth]{assets/f1b.png}
  \end{subfigure}\hfill
  \begin{subfigure}{.4\textwidth}
    \includegraphics[width=\linewidth]{assets/f1c.png}
  \end{subfigure}
  \caption{
    (a) $T_{1}$ optimized gates (red) and analytic gates (blue).
    (b) $T_{1}$ interpolation function used in optimization. Markers
    denote the time-averaged, absolute amplitude of each gate.
    (c) Lindblad master equation simulation with $T_{1}$ dissipation
    for successive gate applications. The reported gate error is computed after
    each gate appliaction.
  }
\end{figure*}

%% S4
\section{Longitudinal Relaxation Awareness}
(Strategy) We seek to minimize the probability
that the qubit decays as a result of longitudinal
relaxation. To this end we add the
longitudinal relaxation probability
to the augmented state vector
\begin{equation}
  P_{1}(t) = \int_{0}^{t} T_{1}^{-1}(a(t^{\prime})) dt^{\prime}
\end{equation}
Setting the target
longitudinal relaxation probability to 0 results in
a quadratic cost at each knot point
of the form ${\lvert P_{1}(t_{k}) \rvert}^{2}$.

$T_{1}(a_{k})$
is obtained at each knot point by evaluating
a spline interpolant fit to
experimentally obtained data of the form
$\{(a, T_{1})\}$.
Calculating $T_{1}$ directly from theoretical
considerations requires many high-rank
eigendecompositions, which
is computationally expensive. Additionally,
$T_{1}$ values are known to fluctuate greatly
with laboratory temperatures \cite{klimov2018fluctuations}.
Interpolating $T_{1}$ from experimental data increases
the fridge truth of the simulation.

For a given $T_{1}$ time, a shorter gate duration would tend to
favor a lower longitudinal relaxation probability.
We allow the optimizer to tune the gate duration by
making the time step between each knot point $\Delta t_{k}$
a decision variable.
Promoting $\Delta t_{k}$ to a decision variable, rather
than the number of knot points $N$, preserves the
Markovian decision structure of the trajectory
optimization problem. To ensure numerical
integration accuracy is maintained we add a bound
constraint at each knot point
$2.5\textrm{e-}3 \ \textrm{ns} \le
\Delta t_{k} \le 1\textrm{e-}2 \ \textrm{ns}$.
Note that this bound constraint is allowed to be
broken for intermediate iterations of the optimization,
so we add the square root of the time step $\sqrt{\Delta t_{k}}$
to the augmented control vector and use the squared root
of the time step $\lvert \Delta t_{k} \rvert$ in optimization.

(Results) The $Y/2$ and $Z/2$ gates converge on similar solutions, a periodic
waveform with amplitude $\sim 0.2 \textrm{GHz}$ (see Figure 1). Both extend their gate times
beyond their analytic counterpart, trading longer gate times for access
to higher amplitudes and therefore higher $T_{1}$ times. Both gates reduce
their probability of longitudinal relaxation per gate by a factor of $\sim 5$ over
their analytic counterparts. Furthermore, they perform simlarly in
the concatenated gate application comparision, maintaining a total gate error under $10^{-2}$
over $10^{3}$ gate applications, for an average gate error on the order of $10^{-5}$.
We define gate error as the infidelity of the final state and the target state,
averaged over multiple pseudo-randomly
generated initial states. The $X/2$ gate

%% T1
%% TODO: maybe for appendix.
\begin{table}[ht]
  \begin{tabular}{c | c | c | c}
    & Analytic  & QOC &\\
    Gate & $P_{1}\ (10^{-5})$ & $P_{1}\ (10^{-5})$ & $P_{1\textrm{A}} / P_{1\textrm{Q}}$\\
    \hline
    Z/2 & 5.745 & 1.149 & 5.000\\
    Y/2 & 5.253 & 1.157 & 4.540\\
    X/2 & 16.251 & 4.488 & 3.621\\
  \end{tabular}
  \caption{Probability of longitudinal relaxation for each gate
    evaluated at the gate's duration.}
\end{table}

%% F2
\begin{figure*}[ht]
  \begin{subfigure}{.315\textwidth}
    \includegraphics[width=\linewidth]{assets/f2a.png}
  \end{subfigure}\hfill
  \begin{subfigure}{.4\textwidth}
    \includegraphics[width=\linewidth]{assets/f2c.png}
  \end{subfigure}\hfill
  \begin{subfigure}{.23\textwidth}
    \includegraphics[width=\linewidth]{assets/f2b.png}
  \end{subfigure}

  \caption{
    (a) X/2 gates robust to qubit frequency detunings constructed using the
    analytic, 2-point sampling, 8-point unscented sampling, and (2\textsuperscript{nd},
    3\textsuperscript{rd})-order derivative methods at a gate duration of $56.8$ns.
    (b) Single gate error as
    a function of the gate duration at a one-percent
    detuning from the nominal qubit frequency for all methods.
    (c) Single gate error as a function of the detuning from the nominal
    qubit frequency at gate durations of $56.8$ns and $120$ns for the analytic
    and 2\textsuperscript{nd}-order derivative methods.
  }
\end{figure*}

%% F3
\begin{figure}[h]
  \begin{subfigure}{\linewidth}
    \includegraphics[width=\linewidth]{assets/f3a.png}
  \end{subfigure}

@@

  \caption{
    (a) $X/2$ gates robust to flux offsets constructed using the analytic,
    (2, 4)-point sampling, and (2\textsuperscript{nd}, 3\textsuperscript{rd})-order
    derivative methods at a gate duration
    of $56.8$ns. (b) Simulation of stochastic 1/$f$ flux noise for
    successive gate applications. The reported gate error is measured
    after each gate application.
  }
\end{figure}

%% S5
\section{Robustness to Parameter Deviations}
(Strategy) Magentic flux noise,
Johnson-Nyquist current noise present in the flux bias
lines, and temperature dependent gain fluctuations
may cause the flux amplitude to deviate from its
desired value by an amount $\delta a$. Additionally,
residual calibration errors and fluctuations may cause
the qubit frequency to deviate from its measured value
by an amount $\delta f_{q}$.

System parameter
deviations are typically addressed with dynamic decoupling
sequences \cite{merrill2014progress},
the DRAG scheme \cite{krantz2019quantum}, or
geometric phase considerations
\cite{xu2020nonadiabatic} \cite{han2020experimental}.
Dynamic decoupling sequences compose rotations on the
Bloch sphere so that erroneous rotations arising due
to the parameter deviation are cancelled.
%% The latter two methods are inflexible.
We draw on the
robust control literature to demonstrate
numerical techniques for engineering robustness
to parameter deviations.

We propose two methods for engineering robustness
to parameter deviations. The first is the
derivative method. We draw on the intuition that
the sensitivity of the state evolution to the parameter
$\lambda$ is encoded in the derivative of
the state with respect to that parameter
$\partial_{\lambda}^{l} \ket{\psi}$. The derivative
scheme aims to minimize the norm of this derivative.
The derviatives of the state with respect to the parameter
of interest are propgated in the
augmented state vector. Their dynamics are found by differentiating
the TDSE dynamics with respect to the parameter of interest
(see Appendix). Setting the target derivatives of the state
to the zero vector results in quadratic costs at each
knot point
${\lvert \partial_{\lambda}^{l} \ket{\psi} \rvert}^{2}$.
In the $l$\textsuperscript{th}-order derivative method, the norm of
all derivatives of order less than or equal to $l$ are penalized. For example,
in the 2\textsuperscript{nd}-order derivative method both the
1\textsuperscript{st}-order state derivative
(${\lvert \partial_{\lambda} \ket{\psi_{k}} \rvert}^{2}$)
and the 2\textsuperscript{nd}-order state derivative
(${\lvert \partial^{2}_{\lambda} \ket{\psi_{k}} \rvert}^{2}$)
contribute quadratic costs at each knot point.

The second scheme we analyze is the sampling method,
which is well studied in the robust control literature
\cite{manchester2016derivative} \cite{tronarp2016sigma}.
We add additional
states to the augmented state vector that follow deviant dynamics, and
penalize the state's difference from the target state.
The deviant dynamics are given by the system
Hamiltonian with the parameter of interest replaced by
a deviant value $\lambda \gets \lambda + \delta \lambda$
The parameter's deviation is typically chosen
to represent the statistical distribution of the parameter.
For example, in the 2-point sampling method we
propagate the additional states $\ket{\psi^{\pm}}$ where
$\lambda^{\pm} = \lambda \pm \sigma_{\lambda}$, with
$\sigma_{\lambda}$ being the standard deviation of $\lambda$.
The difference between the final states and the target
state is penalized, resulting in quadratic costs at
each knot point
${\lvert \ket{\psi^{\pm}_{k}} - \ket{\psi_{f}} \rvert}^{2}$.
For the 4-point sampling method we sample at two standard deviations
above and below the nominal parameter value in addition to one standard
deviation above and below.

(Comparison) To demonstrate the applicability of these techniques
to mitigate system parameter deviations,
we consider the task of achieving a single $X/2$
gate subject to a constant qubit frequency detuning
($f_{q} \gets f_{q} + \delta f_{q}$).
%% We use a compensating for off-resonance error
%% (CORPSE) dynamic decoupling pulse that is optimized
%% to mitigate first-order error arising due to the detuning.
%% The pulse is composed of
%% three rotations about the $\hat{x}$, $-\hat{x}$, and $\hat{x}$ axes
%% of the Bloch sphere
%% designed to compensate for erroneous rotations
%% (see Appendix).
We take $\sigma_{f_{q}} = 1\% f_{q}$ to be one standard devation, and equip
the sampling robustness methods accordingly.
For each method we compute the gate error for one simulated gate application
subject to the stated deviant dynamics.
